---
title: "Getting and Cleaning Data Code Book"
author: "Kelcy Klein"
date: "`r Sys.Date()`"
output: html_document
---
## This is the accompanying codebook for the Getting and Cleaning Data Week 4 Project

## 1. The neccesary packages and dataset is downloaded, unzipped, and extracted into the working directory
```{r download_data, include=FALSE}
packages <- c("data.table", "reshape2", "dplyr")
sapply(packages, require, character.only = TRUE, quietly = TRUE)
path <- getwd()
url <- "https://d396qusza40orc.cloudfront.net/
        getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(url, file.path(path, "dataFiles.zip"))
unzip(zipfile = "dataFiles.zip")
```

## activities and features.txt table are read into actvities and features df based on category fxn andactivities


```{r features and activities}
features <- read.table("UCI HAR Dataset/features.txt", 
                       col.names = c("n", "functions"))
activities <- read.table("UCI HAR Dataset/activity_labels.txt", 
                         col.names = c("code", "activity"))
```

## Train and test datasets imported based on fxn and code that corresponds w/ their activity

```{r train and test}
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt", 
                           col.names = "subject")
x_test <- read.table("UCI HAR Dataset/test/x_test.txt",
                     col.names = features$functions)
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", 
                     col.names = "code")
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", 
                            col.names = "subject")
x_train <- read.table("UCI HAR Dataset/train/x_train.txt", 
                      col.names = features$functions)
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", 
                      col.names = "code")
```

## Tables are merged and bound using rbind, cbind

```{r rbind and cbind}
x <- rbind(x_train, x_test)
y <- rbind(y_train, y_test)
subject <- rbind(subject_train, subject_test)
merge_data <- cbind(subject, y, x)
```

## 2. Data is cleans and essential data of code/activity, the mean, and std dev are selected into a new dataframe "tinydata"

```{r tindydata}
tidydata <- merge_data %>% select(subject, code, contains("mean"), contains("std"))
```

## 3. Column names of tidy dataset are renamed to make them more meaningful and consistent

```{r gsub}
tidydata$code <- activities[tidydata$code, 2]
names(tidydata) [2] = "activity"
names(tidydata) <- gsub("Acc", "Accelerometer", names(tidydata))
names(tidydata) <- gsub("Gyro", "Gyroscope", names(tidydata))
names(tidydata) <- gsub("BodyBody", "Body", names(tidydata))
names(tidydata) <- gsub("Mag", "Magnitude", names(tidydata))
names(tidydata) <- gsub("^t", "Time", names(tidydata))
names(tidydata) <- gsub("^f", "Frequency", names(tidydata))
names(tidydata) <- gsub("tBody", "TimeBody", names(tidydata))
names(tidydata) <- gsub("-mean()", "Mean", names(tidydata), ignore.case = TRUE)
names(tidydata) <- gsub("-std()", "STD", names(tidydata), ignore.case = TRUE)
names(tidydata) <- gsub("-freq()", "Frequency", names(tidydata), ignore.case = TRUE)
names(tidydata) <- gsub("angle", "Angle", names(tidydata))
names(tidydata) <- gsub("gravity", "Gravity", names(tidydata))
```

## 4. The final dataset is plotted with the mean of activities group by subject into a new text file called Tidy_Data.txt

```{r Tidy_Data.txt}
Tidy_Data <- tidydata %>%
        group_by(subject, activity) %>%
        summarise_all(funs(mean))
write.table(Tidy_Data, "Tidy_Data.txt", row.names = FALSE)
```